---
title: "Qualitative analysis of weight lifting exercises"
author: "Filipe Lima"
date: "24/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as *Jawbone Up*, *Nike FuelBand*, and *Fitbit* it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: [HAR Project](http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).


## Data cleaning

First we will find the most relevant variables in the [Training data](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) for the project. We chose the ones that have the results for roll, yaw, pitch, gyros, accel and magnet for belt, forearm, arm and dumbell, and of course, the classe variable. Then we will create a new dataset using those variables and classe as a factor.

```{r, eval=FALSE}
library(caret)
WLEdataframe <- read.csv("pml-training.csv")
WLEdataframe$classe <- as.factor(WLEdataframe$classe)
#The index of the variables we want
Selection1 <- c(8:10,37:48,60:68,84:86,113:124,151:160)
WLEdataframe <- WLEdataframe[,Selection1]
```

## Splitting and training

We split the dataset using 70% for training. We use cross validation as a method of train control and 10 as the number of resampling iterations. Then, we trained the data using the Random Forest algorithm.

```{r, eval=FALSE}
inTraining <- createDataPartition(y=WLEdataframe$classe, p = .70, list=FALSE)
training <- WLEdataframe[inTraining,]
testing <- WLEdataframe[-inTraining,]
fitControl <- trainControl(method = "cv",number = 10)
modRF <- train(training[,-49],training[,49],data=training,method="rf",
               trControl = fitControl)
modRF
```


We got the following results:

Random Forest 

13737 samples
   48 predictor
    5 classes: 'A', 'B', 'C', 'D', 'E' 

No pre-processing
Resampling: Cross-Validated (10 fold) 
Summary of sample sizes: 12363, 12362, 12364, 12363, 12364, 12365, ... 
Resampling results across tuning parameters:

| mtry | Accuracy  | Kappa     |
|------|-----------|-----------|
| 2    | 0.9918467 | 0.9896856 |
| 25   | 0.9916290 | 0.9894103 |
| 48   | 0.9861689 | 0.9825029 |

Accuracy was used to select the optimal model using the largest value.
The final value used for the model was mtry = 2.

## Testing and Results

Now, we will test the model in the testing data and look at the confusion matrix.

```{r,eval=FALSE}
predictions <- predict(modRF, newdata=testing)
confusionMatrix(predictions, testing$classe)
```

Confusion Matrix and Statistics

| Prediction | A      | B    | C    | D   | E    |
|------------|--------|------|------|-----|------|
| A          | 1672   | 13   | 0    | 0   | 0    |
| B          | 2 1123 | 1123 | 9    | 0   | 0    |
| C          | 0      | 3    | 1016 | 23  | 1    |
| D          | 0      | 0    | 1    | 941 | 4    |
| E          | 0      | 0    | 0    | 0   | 1077 |

Overall Statistics
                                          
               Accuracy : 0.9905          
                 95% CI : (0.9877, 0.9928)
    No Information Rate : 0.2845          
    P-Value [Acc > NIR] : < 2.2e-16       
                                          
                  Kappa : 0.988           
                                          
Mcnemar's Test P-Value : NA              

Statistics by Class:

|                      | Class A | Class B | Class C | Class D | Class E |
|----------------------|---------|---------|---------|---------|---------|
| Sensitivity          | 0.9988  | 0.9860  | 0.9903  | 0.9761  | 0.9954  |
| Specificity          | 0.9969  | 0.9977  | 0.9944  | 0.9990  | 1.0000  |
| Pos Pred Value       | 0.9923  | 0.9903  | 0.9741  | 0.9947  | 1.0000  |
| Neg Pred Value       | 0.9995  | 0.9966  | 0.9979  | 0.9953  | 0.9990  |
| Prevalence           | 0.2845  | 0.1935  | 0.1743  | 0.1638  | 0.1839  |
| Detection Rate       | 0.2841  | 0.1908  | 0.1726  | 0.1599  | 0.1830  |
| Detection Prevalence | 0.2863  | 0.1927  | 0.1772  | 0.1607  | 0.1830  |
| Balanced Accuracy    | 0.9979  | 0.9918  | 0.9923  | 0.9876  | 0.997   |

We achieved a 99.05% accuracy in this model. 

## The Course Project Prediction Quiz

Now, we will use our model to predict the classes for this [Testing Dataset](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv), with 20 observations.

```{r,eval=FALSE}
predictDATAFRAME <- read.csv("pml-testing.csv")
predict(modRF, newdata=predictDATAFRAME)
```

B A B A A E D B A A B C B A E E A B B B
Levels: A B C D E

This result is 100% correct.